---
title: "Impact of Marine Protected Areas on California Spiny Lobster distribution along Santa Barbara coast"
subtitle: "A report prepared for ESM 206"
author: "Alicia Fennell and Rachel Rhodes"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

# Introduction 

Starting in 2012, data has been collected on the abundance, size, and location of the California spiny lobster (Panulirus interruptus) in an effort to understand the ecology of the coastal kelp forests along the Santa Barbara coast. California spiny lobsters are an important predator species that keep kelp-eating species like sea urchin in check. 


# Data and Methods

Santa Barbara Coastal Long Term Ecological Research program at UC Santa Barbaraca collected data for the abundance, size, and location of California spiny lobster along the coast of the Santa Barbara channel from 2012-2018. Spiny lobster abundance and size data was collected during annual dives before the start of the fishing season in late summer at five sites.Divers recorded the number and size of spiny lobster at four 60 x 5 m (300 m2 total) transects at each site. Two of the sites, Naples and Isla Vista, are Marine Protected Areas that were established in January 2012. Mean lobster abundance for Marine Protected Area sites and non-protected sites were compared by two-sample t-tests. All analysis and figured were prepared using R software version 3.6.1. 


```{r}

# Attach packages

library(tidyverse)
library(here)
library(janitor)
library(tidyr)
library(ggridges)
library(directlabels)
library(RColorBrewer)
library(effsize)
library(kableExtra)

```


```{r}

## Read in data and make it tidy

lobster_ab <- read_csv("lobster_abundance_sbc_lter.csv", na = "-99999") %>% 
  clean_names()

lobster_tidy <- lobster_ab %>% 
  uncount(count) %>% 
  mutate(site = replace(site, site == "IVEE", "Isla Vista")) %>%
  mutate(site = replace(site, site == "CARP", "Carpinteria")) %>% 
  mutate(site = replace(site, site == "AQUE", "Arroyo Quemado")) %>% 
  mutate(site = replace(site, site == "MOHK", "Mohawk")) %>% 
  mutate(site = replace(site, site == "NAPL", "Naples")) %>% 
  mutate(site_type = if_else(site %in% c("Isla Vista","Naples"), "MPA", "Non-MPA"))

## Make dataset for Results A to see abundance at differnent sites through the years

annual_ab_site <- lobster_tidy %>% 
  group_by(year) %>% 
  count(site) %>% 
  mutate(site_type = if_else(site %in% c("Isla Vista","Naples"), "MPA", "Non-MPA"))

 ## maybe delete this? Possibly incorrect summary table of mean sizes of MPA vs non-MPA sites in 2012 and 2018

lobster_summaries <- lobster_tidy %>% 
  filter(year == c(2012, 2018)) %>% 
  group_by(site_type, year) %>% 
  summarize(mean_size = mean(size_mm, na.rm = TRUE))

## Make a dataset for just 2012
  
year_2012 <- lobster_tidy %>% 
  filter(year == "2012")

## Make a dataset for just 2018

year_2018 <- lobster_tidy %>% 
  filter(year == "2018")
         
```


```{r}

## Exploratory graphs

# ggplot(annual_ab_site, aes(x = year, y = n)) +
#   geom_line(aes(color = site)) +
#   theme_minimal()
# 
# ggplot(annual_ab_site, aes(x = year, y = n)) +
#   geom_bar(stat = "identity") +
#   facet_wrap(~site) +
#   theme_minimal() 

```
## Results

#### 1. Effects of Marine Protected Area designation on spiny lobster abundance



```{r}

## Results A - visualalization showing  changes in annual lobster abundace at the five sites over time, designating MPA versus non-MPA sites. 

ggplot(annual_ab_site, aes(x = year, y = n)) +
  geom_line(aes(group = site,
                color = site_type), 
            size = 0.75) +
  theme_minimal()+
  theme(
    legend.position = c(0.25, .88),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    legend.background = element_rect("white"))+
  xlim(2012,2019.5) +
  geom_dl(aes(label=site,
              color = site_type),
          method = "last.points") +
  ggtitle("CALIFORNIA SPINY LOBSTER ABUNDANCE",
          subtitle = "Santa Barbara Channel (2012 - 2018)")+
  labs(x = "Year",
       y = "Lobster count",
       color = "Site Type")
```
***Figure 1*** *California spiny lobster abundance at sites within Marine Protected Area  and sites not within  Marine Protected Areas between 2012 and 2018. Between the years 2017 and 2018, lobster populations at the marine protected sites grew, while populations at the non-marine protected sites declined or remained the same.*


#### 2. Lobster size distribution shifts between 2012 and 2018

```{r}
## Data frame for exploring lobster size distribution shifts in 2012 and 2018

lobster_size <- lobster_tidy %>% 
  filter(year == c(2012, 2018)) %>% 
  mutate(site = fct_relevel(site, c("Isla Vista", "Naples", "Arroyo Quemado", "Carpinteria", "Mohak")))


```

```{r}
## Results B - Visualization of size distribution for lobsters obvserved at the 5 different sites in 2012 and 2018. Graph shows the community size distributions at each in the 2 years. 

# ggplot(lobster_size, aes(x = site, y = size_mm, fill = factor(year))) +
#   geom_violin()+
#   stat_summary(fun.y = "mean", geom = "point", size = 3)+
#   scale_fill_brewer()+
#   labs(x = "Site",
#        y = "Size (mm)",
#        fill = "Year")+
#   ggtitle("CALIFORNIA SPINY LOBSTER DISTRIBUTION 2012 & 2018",
#           subtitle = "Santa Barbara Channel")+
#   theme_minimal()+
#   theme(legend.position = c(0.78, .88),
#         legend.justification = c("left", "top"),
#         legend.box.just = "right",
#         legend.margin = margin(6, 6, 6, 6),
#         legend.background = element_rect("white"))

```
```{r}

## Jitterplot not really working- unable to separate the two years the way we want

ggplot(lobster_size, aes(x = factor(year), y = size_mm)) +
  geom_jitter(aes(color = factor(year)), 
              alpha = 0.5)+
  geom_boxplot(width = 0.3)+
  facet_grid(~site) +
  labs(x = "Site",
       y = "Size (mm)",
       fill = "Year")+
  ggtitle("CALIFORNIA SPINY LOBSTER DISTRIBUTION 2012 & 2018",
          subtitle = "Santa Barbara Channel")+
  theme_minimal()
  # theme(legend.position = c(0.78, .88),
  #       legend.justification = c("left", "top"),
  #       legend.box.just = "right",
  #       legend.margin = margin(6, 6, 6, 6),
  #       legend.background = element_rect("white"))


```
***Figure 2*** *Size distribution of California spiny lobster at each side in 2012 compared to 2018. The abundance of spiny lobster in ...* 

#### 3. Mean lobster sizes between site types (MPA and non-MPA) in 2012 and 2018


-----
```{r, include = FALSE }



## Starting t tests, create 4 vectors with lists of sizes


MPA_2012 <- year_2012 %>% 
  filter(site_type == "MPA") %>% 
  pull(size_mm)

non_MPA_2012 <- year_2012 %>% 
  filter(site_type == "Non-MPA") %>% 
  pull(size_mm)

MPA_2018 <- year_2018 %>% 
  filter(site_type == "MPA") %>% 
  pull(size_mm)

non_MPA_2018 <- year_2018 %>% 
  filter(site_type == "Non-MPA") %>% 
  pull(size_mm)

## #1 size differences in 2012 between MPA and non-MPA

size_2012_ttest <- t.test(MPA_2012, non_MPA_2012)

size_2012_ttest

#Cohen's d

size_2012_d <- effsize::cohen.d(MPA_2012, non_MPA_2012)

size_2012_d

## Summary statistics

n_mpa_2012 <- as.numeric(length(MPA_2012))
n_nonMPA_2012 <- as.numeric(length(non_MPA_2012))

mean_MPA_2012 <- mean(MPA_2012, na.rm = TRUE)
sd_MPA_2012 <- sd(MPA_2012, na.rm = TRUE)

mean_nonMPA_2012 <- mean(non_MPA_2012, na.rm = TRUE)
sd_nonMPA_2012 <- sd(non_MPA_2012, na.rm = TRUE)

median_mpa_2012 <- median(MPA_2012, na.rm = TRUE)
median_mpa_2018 <- median(MPA_2018, na.rm = TRUE)

median_nonmpa_2012 <- median(non_MPA_2012, na.rm = TRUE)
median_nonmpa_2018 <- median(non_MPA_2018, na.rm = TRUE)

```

Is there a significant difference in lobster size between MPA and non-MPA sites in 2012? 

Mean lobster size (mm) in MPA sites (`r round(mean_MPA_2012, 2)` $\pm$ `r round(sd_MPA_2012, 2)`, n = `r n_mpa_2012`) differed significantly from the mean lobster size in non-MPA sites (`r round(mean_nonMPA_2012, 2)` $\pm$ `r round(sd_nonMPA_2012, 2)`, n = `r n_nonMPA_2012`) by using a two-sample t-test (t(`r round(size_2012_ttest$parameter, 2)`) = `r round(size_2012_ttest$statistic, 2)`, *p* = `r round(size_2012_ttest$p.value, 3)`). The effect size was medium (Cohen's d = `r round(size_2012_d$estimate, 2)`).


```{r, include = FALSE }
## #2 size differences in 2018 between MPA and non-MPA


size_2018_ttest <- t.test(MPA_2018, non_MPA_2018)

size_2018_ttest

#Cohen's d

size_2018_d <- effsize::cohen.d(MPA_2018, non_MPA_2018)

size_2018_d

n_mpa_2018 <- as.numeric(length(MPA_2018))
n_nonMPA_2018 <- as.numeric(length(non_MPA_2018))

mean_MPA_2018 <- mean(MPA_2018, na.rm = TRUE)
sd_MPA_2018 <- sd(MPA_2018, na.rm = TRUE)

mean_nonMPA_2018 <- mean(non_MPA_2018, na.rm = TRUE)
sd_nonMPA_2018 <- sd(non_MPA_2018, na.rm = TRUE)

```

Is there a significant difference in lobster size between MPA and non-MPA sites in 2018? 

Mean lobster size (mm) in MPA sites (`r round(mean_MPA_2018, 2)` $\pm$ `r round(sd_MPA_2018, 2)`, n = `r n_mpa_2018`) differed significantly from the mean lobster size in non-MPA sites (`r round(mean_nonMPA_2018, 2)` $\pm$ `r round(sd_nonMPA_2018, 2)`, n = `r n_nonMPA_2018`) by using a two-sample t-test (t(`r round(size_2018_ttest$parameter, 2)`) = `r round(size_2018_ttest$statistic, 2)`, *p* < .001). The effect size was medium-small (Cohen's d = `r round(size_2018_d$estimate, 2)`).


```{r, include = FALSE }
## #3 size differences in MPA between 2012 and 2018

MPA_diff_ttest <- t.test(MPA_2012, MPA_2018)

MPA_diff_ttest

#Cohen's d

MPA_diff_d <- effsize::cohen.d(MPA_2012, MPA_2018)

MPA_diff_d

```

Is there a significant difference in lobster size between 2012 and 2018 in MPA sites? 

Mean lobster size (mm) in MPA sites in 2012 (`r round(mean_MPA_2012, 2)` $\pm$ `r round(sd_MPA_2018, 2)`, n = `r n_mpa_2012`) differed significantly from the mean lobster size in MPA sites in 2018 (`r round(mean_MPA_2018, 2)` $\pm$ `r round(sd_MPA_2018, 2)`, n = `r n_mpa_2018`) by using a two-sample t-test (t(`r round(MPA_diff_ttest$parameter, 2)`) = `r round(MPA_diff_ttest$statistic, 2)`, *p* < .001). The effect size was large (Cohen's d = `r round(MPA_diff_d$estimate, 2)`).

```{r, include = FALSE }

## #4 size differences in non-MPA sites between 2012 and 2018

non_MPA_diff_ttest <- t.test(non_MPA_2012, non_MPA_2018)

non_MPA_diff_ttest

#Cohen's d

non_MPA_diff_d <- effsize::cohen.d(non_MPA_2012, non_MPA_2018)

non_MPA_diff_d

```

Is there a significant difference in lobster size between 2012 and 2018 in non-MPA sites? 

Mean lobster size (mm) in non-MPA sites in 2012 (`r round(mean_nonMPA_2012, 2)` $\pm$ `r round(sd_nonMPA_2018, 2)`, n = `r n_nonMPA_2012`) did not differ significantly from the mean lobster size in non-MPA sites in 2018 (`r round(mean_nonMPA_2018, 2)` $\pm$ `r round(sd_nonMPA_2018, 2)`, n = `r n_nonMPA_2018`) by using a two-sample t-test (t(`r round(non_MPA_diff_ttest$parameter, 2)`) = `r round(non_MPA_diff_ttest$statistic, 2)`, *p* = `r round(non_MPA_diff_ttest$p.value, 3)`). The effect size was small (Cohen's d = `r round(non_MPA_diff_d$estimate, 2)`).


```{r}
### table

# Make table subset

table_data <- lobster_tidy %>% 
  select(year, site, size_mm, site_type) %>% 
  group_by(site_type, year) %>% 
  summarize(mean_size = mean(size_mm), sd = sd(size_mm), n = length(size_mm)) %>% 
  filter(year %in% c(2012, 2018))

round_table <- table_data %>% 
  mutate_at(vars(-n, -site_type, -year), funs(round(., 2)))

lobs_table <- round_table %>% 
  kable(col.names = c("Site Type",
                      "Year",
                      "Mean Size (mm)", 
                     "Standard Deviation", 
                     "Sample Size (n)")) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = F,
                position = "left") %>% 
  add_header_above(c("Summary of Lobster Size Distribution" = 5))

lobs_table


```


## Summary


## References 

**Reed D. 2019.** *SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012.* Environmental Data Initiative. https://doi.org/10.6073/pasta/a593a675d644fdefb736750b291579a0. Dataset accessed 11/15/2019.